---
title: "Identifying Those Tricky Little Micro-moths"
author: Paul J. Palmer
output: 
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    slide_level: 1
    includes:
          in_header: "preamble.tex" 
    keep_tex:  true
---

```{r child = "_header.Rmd"}
```

# What Is A Micro-moth?
-   This is a tricky question to answer as there no taxonomic difference, as in butterflies and moths. Even that is a bit blurred.
-   So far as I can tell, the larger moths are those that were described in South vols 1 and 2, and the rest are the micro-moths.
-   In the olden days when we used a light trap there we so many large moths that we were never tempted by the micro's that were trampled by the Noctua pronuba (Large Yellow Underwing) which came in their hundreds.
-   Micro's were collected by rearing out series collected off food plants and mines.
-   The first (and current) editions of "The Field Guide To The Smaller Lepidoptera" contains no illustrations of adult moths at all.

# Wot! No Illustrations?
Don't worry, I will use them later in this presentation.


-   Without illustrations the emphasis for identification rested on descriptions of the larval behaviour and food plants plus the overall appearance of the adult being consistent with the overall family and genus of the specimen.
-   Just to emphasise, you had to be really proficient with your plant identification too.
-   My old copy of Clapham, Tutin, and Warburgh "The Excursion Flora Of The British Isle" was also without images.
-   To those of you only used to modern illustrated field guides this must sound very strange, but the point I make is that a good identification is based on a collection of evidence which is used to eliminate all but one possibility.

# First Steps
-   
-   
-   
-   The availability of 'Apps' that seek to make an identification from such photographs, especially when they work on a common species, reinforces this approach.


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "First Steps"
bullets.text <- list(
"Light traps catch fewer moths these days, so the micro-moths are more obvious.",
"The publication of the excellent illustrated 'Field Guide to the Micro-Moths of Great Britain and Ireland' has helped popularise this group further.",
"BUT, the temptation for beginners is to 'pattern match' from a field guide.",
"The availability of 'Apps' reinforces this approach."
  
)
supporting.image <- "micro-field-guide.jpg"
supporting.text <- "Field Guide to the Micro Moths"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Identification Is A Process"
bullets.text <- list(
"Examine your specimen closely.",
"Treat ID as process and start by writing down everything that you know and can see.",
"Use these notes to try and work out the Family.",
"There is a good visual key to families in 'Field Guide to the Micro-Moths of Great Britain and Ireland'.",
"Lepidoptera keys to species level are few and far between."
  
)
supporting.image <- "micro-lep-key.jpg"
supporting.text <- "A key to the Micro Moths"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Candidates For ID"
bullets.text <- list(
"This is a tedious and time-consuming process.",
"You should now have a list that includes: size, locality and type of habitat; date; and other obvious features.",
"A photograph is also helpful - look at the posture and position of the antennae."
  
)
supporting.image <- "Pcontaminella-PML-2021.jpg"
supporting.text <- "What can we see here?"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```




```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Getting To The Family"
bullets.text <- list(
"The antennae are laying flush along the body",
"This is a member of the Pyraloidea family.",
"The face has a distinctive 'nose' and face:",
"So this is a member of the Crambinae subfamily, which are colloquially known as 'Grass moths'. ",
"*We have reduced the number of candidate species to about 17*."
  
)
supporting.image <- "Pcontaminella-PML-2021.jpg"
supporting.text <- "More  text"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "The Answer Must Be In This Book"
bullets.text <- list(
"Text",
"Text.",
"Text."
  
)
supporting.image <- "placeholder.jpg"
supporting.text <- "More  text"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```



```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Now we know it is a Cranbid (Grass Moth)"
bullets.text <- list(
"Date: 2021-07-25, VC55; Recorder: Pete Leonard; Size: FL 12 mm.",
"Could this be *Pediasia contaminella* which is rare in VC55?",
"This specimen lacks the cross lines the dark point in the discal region typical of *Pediasia contaminella*.",
"But *Pediasia contaminella* said to have a distinctive upwards resting posture."
  
)
supporting.image <- "PJP20220218-0001a.jpg"
supporting.text <- "The set moth is not a perfect match for any of the 17 species. "

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```
Specimens seen in Sussex by PJP were all well marked and instantly recognisable in both posture and appearance.

