---
title: "Identifying Those Tricky Little Micro-moths"
author: Paul J. Palmer
output: 
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    slide_level: 1
    includes:
          in_header: "preamble.tex" 
    keep_tex:  true
---

```{r child = "_header.Rmd"}
```

# What Is A Micro-moth?
-   This is a tricky question to answer as there no taxonomic difference, as in butterflies and moths. Even that is a bit blurred.
-   So far as I can tell, the larger moths are those that were described in South vols 1 and 2, and the rest are the micro-moths.
-   In the olden days when we used a light trap there we so many large moths that we were never tempted by the micro's that were trampled by the Noctua pronuba (Large Yellow Underwing) which came in their hundreds.
-   Micro's were collected by rearing out series collected off food plants and mines.
-   The first (and current) editions of "The Field Guide To The Smaller Lepidoptera" contains no illustrations of adult moths at all.

# Wot! No Illustrations?
Don't worry, I will use them later in this presentation.


-   Without illustrations the emphasis for identification rested on descriptions of the larval behaviour and food plants plus the overall appearance of the adult being consistent with the overall family and genus of the specimen.
-   Just to emphasise, you had to be really proficient with your plant identification too.
-   My old copy of Clapham, Tutin, and Warburgh "The Excursion Flora Of The British Isle" was also without images.
-   To those of you only used to modern illustrated field guides this must sound very strange, but the point I make is that a good identification is based on a collection of evidence which is used to eliminate all but one possibility.



```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "First Steps"
bullets.text <- list(
"Light traps catch fewer moths these days, so the micro-moths are more obvious.",
"The publication of the excellent illustrated 'Field Guide to the Micro-Moths of Great Britain and Ireland' has helped popularise this group further.",
"BUT, the temptation for beginners is to 'pattern match' from a field guide.",
"The availability of 'Apps' reinforces this approach."
  
)
supporting.image <- "micro-field-guide.jpg"
supporting.text <- "Field Guide to the Micro Moths"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Identification Is A Process"
bullets.text <- list(
"Examine your specimen closely.",
"Treat ID as process and start by writing down everything that you know and can see.",
"Use these notes to try and work out the Family.",
"There is a good visual key to families in 'Field Guide to the Micro-Moths of Great Britain and Ireland'.",
"Lepidoptera keys to species level are few and far between."
  
)
supporting.image <- "micro-lep-key.jpg"
supporting.text <- "A key to the Micro Moths"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "A Tricky Moth For ID"
bullets.text <- list(
  "Making a list of candidate species is a tedious and time-consuming process.",
"You should now have a list that includes: size, locality and type of habitat; date; and other obvious features.",
"A photograph is also helpful - look at the posture and position of the antennae."
  
)
supporting.image <- "Pcontaminella-PML-2021.jpg"
supporting.text <- "What can we see here?"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```

We will illustrate the ID process by following one moth tricky moth through its journey.


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Checking The Basics"
bullets.text <- list(
"First check that our specimen is a moth.",
"This is not as daft as it sounds as it is not unheard of for a Caddis (Order: Trichoptera) to be queried as a moth.",
"In this case the overall appearance is definitely Lepidoptera."

  
)
supporting.image <- "order.jpg"
supporting.text <- "Tree of life (biology)"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Checking External Features"
bullets.text <- list(

"All the short jawed Lepidoptera are micro moths, so it is quite likely that you will come across them, especially those that might be tricky to ID.",
"This specimen has a long coiled tongue, rather than short jaws, as have the majority of the Lepidoptera.",
"The long long coiled tongue means that this moth is a member of the **Glossata**."

  
)
supporting.image <- "suborder.jpg"
supporting.text <- "Tree of life (biology)"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```




```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Getting To The Family"
bullets.text <- list(
  "The face has a distinctive 'nose' and face. Also the antennae are laying flush along the body:",
  "This is a member of the **Pyraloidea** super-family."
  
)
supporting.image <- "Pcontaminella-PML-2021.jpg"
supporting.text <- "Note the posture."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Following Branches Of The Tree"
bullets.text <- list(
"Now, the **Crambidae** family are very variable, but this moth is a member of the subfamily **Crambinae**, as the long wings can wrap tightly around the body. Colloquially these are known as 'Grass moths'.",
"*We have reduced the number of candidate species to about 17*."
  
  
)
supporting.image <- "subfamily.jpg"
supporting.text <- "*The Natural History Museum futher subdivide this group into tribes, but these do not seem to be in general use*."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```




```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Confirming the ID"
bullets.text <- list(
"We have gone as far as we can with external features.",
"The way forward is examination of the genitalia - **GenDet**.",
"We start by relaxing the dead moth"
  
)
supporting.image <- "relaxing.jpg"
supporting.text <- "Relaxing chemicals are water based."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Setting"
bullets.text <- list(
"The relaxed moth is pinned and spread on a setting board.",
"Note how everything is carefully labelled.",
"It is then dried for a couple of weeks.",
"This process allows us to preserve as much of the specimen as possible."
  
)
supporting.image <- "setting_moths.jpg"
supporting.text <- "A group of moths on the setting board."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "We know it is a **Crambinae** (Grass Moth)"
bullets.text <- list(
"Date: 2021-07-25, VC55; Recorder: Pete Leonard; Size: FL 12 mm.",
"Could this be *Pediasia contaminella* which is rare in VC55?",
"This specimen lacks the cross lines the dark point in the discal region typical of *Pediasia contaminella*.",
"But *Pediasia contaminella* said to have a distinctive upwards resting posture."
  
)
supporting.image <- "PJP20220218-0001a.jpg"
supporting.text <- "The set moth does not reveal any useful features for ID."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```
Specimens seen in Sussex by PJP were all well marked and instantly recognisable in both posture and appearance.




```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Clarification"
bullets.text <- list(
"The abdomen is removed from the moth and 'cooked' for 30 minutes at 70 degrees in 10% KOH.",
"A label is place in the tube using paper and ink that can survive this process.",
"A body swap would be a disaster."

  
)
supporting.image <- "clarification.jpg"
supporting.text <- "Equiment used for clarification"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```

This process is loosely called *clarification* as it dissolves the soft tissues. Different authors use the term to describe rather different processes.


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Dissecting"
bullets.text <- list(
"The genitalia are very small so a range of bought and homemade tools are used to dissect out the parts needed.",
"They are also very fragile, so you only get one chance on a moth this size.",
"No pressure when you are doing this for someone else!"
  
)
supporting.image <- "dissecting.jpg"
supporting.text <- "The smaller the dissection the more important the work space."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Mounting The Dissection"
bullets.text <- list(
"The parts are washed in de-ionised water.",
"Soaked in a little aqueous mountant.",
"Placed on a slide and a coverslip placed on top.",
"Finally, a temporary label is glued to the slide"
  
)
supporting.image <- "mounted_slides.jpg"
supporting.text <- "Slides are left for a few days to cure."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Ringing The Slide"
bullets.text <- list(
"Text",
"Text.",
"Text."
  
)
supporting.image <- "slideringing.jpg"
supporting.text <- "More  text"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Examination Under The Microscope"
bullets.text <- list(
"Text",
"Text.",
"Text."
  
)
supporting.image <- "PJP20220218-001-raw-dissection.jpg"
supporting.text <- "Raw image"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "The Finished Dissection"
bullets.text <- list(
"Check the dissection against multiple sources as this specimen is not typical.",
"https://mothdissection.co.uk",
"https://lepiforum.org",
"Ask someone who is familiar with the species."
  
)
supporting.image <- "PJP20220218-001-developed-dissection.jpg"
supporting.text <- "Few keys for dissections."

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```



```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "Checking The ID"
bullets.text <- list(
"The dissection is compared with other close relatives.",
"We try to eliminate related species not on the UK list.",
"In this case we find no other candidates other than: *Pediasia contaminella.*"
  
)
supporting.image <- "species.jpg"
supporting.text <- " "

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "To Recap"
bullets.text <- list(
"Identification of a tricky specimen is a tedious process.",
"Most of the time it turns out to be an unusual form of a common species as moths are very variable in size and wing pattern.",
"Environmental stress can cause size and colour variations.",
"Confirming the ID of something unusual requires the collation of a lot of evidence."
  
)
supporting.image <- "placeholder.jpg"
supporting.text <- "More  text"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```

*But sometimes you get lucky!*

