---
title: "Identifying Those Tricky Little Micro-moths"
author: Paul J. Palmer
output: 
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
    slide_level: 1
    includes:
          in_header: "preamble.tex" 
    keep_tex:  true
---

```{r child = "_header.Rmd"}
```

# What Is A Micro-moth?
-   This is a tricky question to answer as there no taxonomic difference, as in butterflies and moths. Even that is a bit blurred.
-   So far as I can tell, the larger moths are those that were described in South vols 1 and 2, and the rest are the micro-moths.
-   In the olden days when we used a light trap there we so many large moths that we were never tempted by the micro's that were trampled by the Noctua pronuba (Large Yellow Underwing) which came in their hundreds.
-   Micro's were collected by rearing out series collected off food plants and mines.
-   The first (and current) editions of "The Field Guide To The Smaller Lepidoptera" contains no illustrations of adult moths at all.

# Wot! No Illustrations?
Don't worry, I will use them later in this presentation.


-   Without illustrations the emphasis for identification rested on descriptions of the larval behaviour and food plants plus the overall appearance of the adult being consistent with the overall family and genus of the specimen.
-   Just to emphasise, you had to be really proficient with your plant identification too.
-   My old copy of Clapham, Tutin, and Warburgh "The Excursion Flora Of The British Isle" was also without images.
-   To those of you only used to modern illustrated field guides this must sound very strange, but the point I make is that a good identification is based on a collection of evidence which is used to eliminate all but one possibility.

# First Steps
-   
-   
-   
-   The availability of 'Apps' that seek to make an identification from such photographs, especially when they work on a common species, reinforces this approach.


```{r  warning=FALSE, message=FALSE, eval=TRUE}

# Create two column beamer slide
header.text <- "First Steps"
bullets.text <- list(
"Light traps catch fewer moths these days, so the micro-moths are more obvious.",
"The publication of the excellent illustrated 'Field Guide to the Micro-Moths of Great Britain and Ireland' has helped popularise this group further.",
"BUT, the temptation for beginners is to 'pattern match' from a field guide.",
"The availability of 'Apps' reinforces this approach."
  
)
supporting.image <- "micro-field-guide.jpg"
supporting.text <- "Field Guide to the Micro Moths"

beamer.bullets.by.image(Header.text = header.text,
                          heading.level = 1,
                           Bullets.text = bullets.text,
                           Supporting.image = supporting.image,
                           Supporting.text = supporting.text )


```



# Size

-   Micromoths are generally small, with wingspans usually less than 20 mm (0.79 inches).
-   Keep in mind that size alone is not a definitive characteristic, as there are larger moths that might still be considered microlepidoptera.

# Wing Shape and Venation:

-   Pay attention to the shape of the wings and their venation (pattern of veins). Different micromoth species may have distinct wing shapes and vein patterns.
-   Some micromoths have narrow, pointed wings, while others may have broader wings.

# Coloration and Patterns:

-   Observe the coloration and patterns on the wings. Micromoths can have intricate patterns, even though they might be subtle.
-   Look for distinctive markings such as spots, lines, or bands on the wings.

# Antennae:

-   Examine the antennae. The shape and characteristics of the antennae can be useful for identification.
-   Micromoths may have thread-like or feathery antennae.

# Resting Posture:

-   Take note of the moth's resting posture. Some micromoths hold their wings flat, while others may fold their wings around their bodies.

# Habitat and Behavior:

-   Consider the habitat in which you find the moth. Different micromoth species may prefer specific environments.
-   Note the behavior of the moth, such as its flight pattern and feeding habits.

# **Use a Field Guide:**

-   A regional field guide to moths and butterflies can be a valuable resource. These guides often provide images, descriptions, and information on distribution.
-   Online resources and mobile apps dedicated to moth identification can also be helpful.

# Photography:

-   If possible, take clear photographs of the micromoth from various angles. This can be useful for later reference or for seeking assistance from experts or online communities dedicated to moth identification.

# Checklist of Common Micromoth Families:

-   Familiarize yourself with common micromoth families in your region. This can help narrow down possibilities.

Remember that micromoth identification can be challenging, and sometimes it may require microscopic examination of specific features. If you're having difficulty, reaching out to entomologists, local insect groups, or online forums dedicated to moth identification can provide valuable assistance.

# Paul's Notes

-   We are talking about the tricky moths, so flicking through pages of a book and hoping get an ID is not really going to work
-   We need to be systematic
-   Systematic identification
